---
title: "COMP3020 Group Project - Task 2.5 Network Analysis"
author: "Miciella Ngo"
date: "2026-10-05"
subtitle: '"Climate Change" vs "Climate Crisis" on Bluesky'
---

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(stringr)
library(tidyr)
library(igraph)
library(ggplot2)

raw <- readRDS("posts_raw.rds")
```

# 2.5 Network Analysis

## Research question

**Which accounts occupy the most important positions in the Bluesky conversation around “climate change” and “climate crisis”, based on reply relationships?**

This analysis uses a directed reply network. A node represents a Bluesky account. A directed edge from account A to account B means that A posted a reply to a post authored by B.

The graph is directed because a reply has a clear sender and receiver. If A replies to B, this does not imply that B replied to A.

## Construct the reply network

```{r build-reply-network}
# Extract the DID of the author who wrote each collected post
raw$source_did <- sapply(raw$author_data, function(x) {
  if (is.null(x$did)) NA_character_ else x$did
})

# Extract the DID of the author being replied to from the Bluesky parent-post URI
raw$target_did <- str_extract(
  raw$in_reply_to,
  "did:plc:[^/]+"
)

# Keep valid reply relationships only
reply_events <- raw |>
  filter(
    !is.na(source_did),
    !is.na(target_did),
    source_did != target_did
  ) |>
  transmute(
    source_did,
    target_did,
    source_handle = author_handle,
    search_term
  )

# Collapse repeated replies between the same two accounts
reply_edges <- reply_events |>
  count(source_did, target_did, name = "weight")

# Build directed weighted graph
g <- graph_from_data_frame(
  reply_edges,
  directed = TRUE
)

g
```

The reply network contains **`r vcount(g)` nodes**, **`r ecount(g)` unique directed edges**, and **`r nrow(reply_events)` reply events**.

## Create readable account labels

```{r account-labels}
# Build a lookup from DID to handle for accounts that appear as collected-post authors
did_lookup <- raw |>
  transmute(
    did = sapply(author_data, function(x) {
      if (is.null(x$did)) NA_character_ else x$did
    }),
    handle = author_handle
  ) |>
  filter(!is.na(did), !is.na(handle)) |>
  distinct(did, .keep_all = TRUE)

# Assign readable labels where available; otherwise keep the DID
vertex_df <- tibble(did = V(g)$name) |>
  left_join(did_lookup, by = "did") |>
  mutate(label = ifelse(is.na(handle), did, handle))

V(g)$label <- vertex_df$label
```

## Network properties

```{r network-properties}
weak_comp <- components(g, mode = "weak")
strong_comp <- components(g, mode = "strong")

in_deg <- degree(g, mode = "in")
out_deg <- degree(g, mode = "out")
total_deg <- degree(g, mode = "all")

network_summary <- tibble(
  Measure = c(
    "Nodes",
    "Unique directed edges",
    "Reply events",
    "Density",
    "Weakly connected components",
    "Largest weak component",
    "Strongly connected components",
    "Median in-degree",
    "Maximum in-degree",
    "Median out-degree",
    "Maximum out-degree"
  ),
  Value = c(
    vcount(g),
    ecount(g),
    nrow(reply_events),
    round(edge_density(g, loops = FALSE), 5),
    weak_comp$no,
    max(weak_comp$csize),
    strong_comp$no,
    median(in_deg),
    max(in_deg),
    median(out_deg),
    max(out_deg)
  )
)

network_summary
```

The network density is **0.0005**, indicating that only a very small proportion of all possible account-to-account reply connections were observed. The graph contains **`r weak_comp$no` weakly connected components**, and the largest weak component contains **`r max(weak_comp$csize)` nodes**.

## Degree distribution

```{r degree-distribution, fig.width=7, fig.height=4.5}
degree_df <- tibble(
  account = V(g)$label,
  total_degree = as.numeric(total_deg),
  in_degree = as.numeric(in_deg),
  out_degree = as.numeric(out_deg)
)

ggplot(degree_df, aes(x = total_degree)) +
  geom_histogram(binwidth = 1, boundary = 0) +
  labs(
    title = "Degree Distribution of the Bluesky Reply Network",
    x = "Total degree",
    y = "Number of accounts"
  ) +
  theme_minimal()
```

Accounts with high in-degree receive replies from many different accounts, while accounts with high out-degree reply to many different accounts.

## Full network visualisation

```{r full-network, fig.width=9, fig.height=9}
set.seed(3020)

label_cut <- quantile(total_deg, 0.99, na.rm = TRUE)

vertex_labels <- ifelse(
  total_deg >= label_cut,
  V(g)$label,
  NA
)

plot(
  g,
  layout = layout_with_fr(g),
  vertex.size = 2 + 2 * log1p(in_deg),
  vertex.label = vertex_labels,
  vertex.label.cex = 0.55,
  vertex.label.dist = 0.4,
  vertex.color = "grey85",
  vertex.frame.color = "grey50",
  edge.width = 0.3 + log1p(E(g)$weight),
  edge.arrow.size = 0.15,
  edge.color = "grey65",
  main = "Full Bluesky Climate Reply Network"
)
```

The full graph shows the overall structure of reply activity. Node size reflects in-degree, so larger nodes are accounts receiving replies from more distinct accounts. Edge width represents the number of replies from one account to another.

## Justified subgraph: largest weakly connected component

The full graph may contain many disconnected components. To examine the most connected part of the conversation, the largest weakly connected component is analysed separately. This is justified because it contains the largest continuous reply structure in the dataset.

```{r largest-component}
largest_component_id <- which.max(weak_comp$csize)

largest_vertices <- V(g)[weak_comp$membership == largest_component_id]

g_lcc <- induced_subgraph(g, vids = largest_vertices)

cat("Largest component nodes:", vcount(g_lcc), "\n")
cat("Largest component edges:", ecount(g_lcc), "\n")
```

```{r largest-component-plot, fig.width=8, fig.height=8}
set.seed(3020)

lcc_in <- degree(g_lcc, mode = "in")
lcc_total <- degree(g_lcc, mode = "all")

# Only label the 10 most connected accounts
top10_lcc <- names(sort(lcc_total, decreasing = TRUE))[1:10]

lcc_labels <- ifelse(
  V(g_lcc)$name %in% top10_lcc,
  V(g_lcc)$label,
  NA
)

plot(
  g_lcc,
  layout = layout_with_fr(g_lcc),
  vertex.size = 4 + 3 * log1p(lcc_in),
  vertex.label = lcc_labels,
  vertex.label.cex = 0.7,
  vertex.color = "grey85",
  vertex.frame.color = "grey50",
  edge.width = 0.5 + log1p(E(g_lcc)$weight),
  edge.arrow.size = 0.2,
  edge.color = "grey60",
  main = "Largest Connected Reply Community"
)
```

The largest component contains **`r vcount(g_lcc)` nodes** and **`r ecount(g_lcc)` edges**. This subgraph is easier to interpret because it focuses on the most interconnected reply structure rather than isolated or very small components.

## Centrality analysis

Two main centrality measures are used:

1.  **In-degree centrality**: counts how many different accounts reply to a given account. A high in-degree indicates that the account attracts attention or discussion.
2.  **Betweenness centrality**: measures how often an account lies on shortest paths between other accounts. A high value suggests that the account occupies a bridging position between different parts of the reply network.

PageRank is also included as a supporting measure.

```{r centrality}
cent <- tibble(
  did = V(g)$name,
  account = V(g)$label,
  in_degree = degree(g, mode = "in"),
  out_degree = degree(g, mode = "out"),
  betweenness = betweenness(
  g,
  directed = TRUE,
  weights = NA,
  normalized = TRUE
),
  pagerank = page_rank(
    g,
    directed = TRUE,
    weights = E(g)$weight
  )$vector
)

top_in <- cent |>
  arrange(desc(in_degree), desc(pagerank)) |>
  slice_head(n = 10)

top_between <- cent |>
  arrange(desc(betweenness)) |>
  slice_head(n = 10)

top_pagerank <- cent |>
  arrange(desc(pagerank)) |>
  slice_head(n = 10)

top_in
top_between
top_pagerank
```

## Top accounts by in-degree

```{r top-indegree-table}
top_in |>
  select(account, in_degree, out_degree, betweenness, pagerank)
```

The account with the highest in-degree is **`r top_in$account[1]`**, with an in-degree of **`r top_in$in_degree[1]`**. This means it received replies from more distinct accounts than any other account in the collected reply network.

## Top accounts by betweenness

```{r top-betweenness-table}
top_between |>
  select(account, in_degree, out_degree, betweenness, pagerank)
```

The account with the highest betweenness centrality is **`r top_between$account[1]`**. This suggests that it occupies an important bridging position between different parts of the network.

## Top accounts by PageRank

```{r top-pagerank-table}
top_pagerank |>
  select(account, in_degree, out_degree, betweenness, pagerank)
```

PageRank highlights accounts that receive replies from structurally important accounts, not just a large number of accounts.

## Framing comparison

To connect the network analysis to the overall project question, reply activity is compared across the two search framings.

```{r framing-comparison}
framing_summary <- reply_events |>
  count(search_term, name = "reply_events") |>
  mutate(
    proportion = reply_events / sum(reply_events)
  )

framing_summary
```

```{r framing-chart, fig.width=6, fig.height=4}
ggplot(framing_summary, aes(x = search_term, y = reply_events, fill = search_term)) +
  geom_col(show.legend = FALSE) +
  labs(
    title = "Reply Events by Search Framing",
    x = "Search framing",
    y = "Number of reply events"
  ) +
  theme_minimal()
```

Of the 816 usable reply events, 538 (65.9%) occurred in posts collected using the “climate change” search term and 278 (34.1%) occurred in the “climate crisis” group. This describes the distribution of reply activity in the collected sample but does not establish that one framing generates greater engagement.

## Interpretation

The reply network provides a direct view of interaction structure within the climate discussion. In-degree identifies accounts that attract replies from multiple other accounts, while betweenness identifies accounts that connect otherwise separate parts of the network.

The largest weakly connected component contains 92 accounts and represents the most interconnected reply community in the dataset. Although the overall network is highly fragmented, this component reveals a clearer local structure of interaction, with several accounts receiving replies from multiple users and acting as focal points within the discussion.

The wider network contained 488 weakly connected components, indicating that the Bluesky climate conversation in this sample was largely fragmented into many small reply groups rather than forming one unified discussion.

The framing comparison connects the network analysis to the broader project question by showing how reply activity was distributed across the two search groups. However, because the groups contained different numbers of collected posts, these raw reply counts should not be interpreted as evidence that one framing generates greater engagement.

## Limitations

Several limitations affect this analysis.

First, the data was collected through Bluesky search for the phrases “climate change” and “climate crisis”, so the network does not represent all climate-related discussion on Bluesky.

Second, the data was collected at one point in time, and Bluesky search results can change. Repeating the same collection process may return different posts.

Third, the target account in each reply is identified by extracting the DID embedded in the `in_reply_to` URI. This is a genuine Bluesky relationship, but not every target account is also present as a collected post author. Therefore, some target nodes cannot be mapped back to readable handles and are displayed as DIDs.

Fourth, the network is based only on reply relationships. It does not include other forms of interaction such as follows, reposts, quotes, or mentions. During network construction, self-replies were excluded and repeated replies between the same pair of accounts were collapsed into a single weighted edge. This avoids inflating the number of unique relationships, but means the edge count represents unique account-to-account connections rather than individual reply events.

Finally, centrality scores describe structural importance within this collected sample only. They should not be interpreted as general influence across the whole Bluesky platform.
