---
title: "2.1 Data Collection"
author: "Aakash Jadhav"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
# Hide package messages and warnings in the output (applies to the whole document)
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

# 2.1 Data Collection

**Topic:** We investigate how the framing of climate discussions on Bluesky differs between posts that say "climate crisis" and posts that say "climate change".

**What?**
We collected public Bluesky posts containing each phrase, then cleaned them so every post belongs to exactly one framing group.

**Why?**
Bluesky is an open social platform with a free public API. Its posts include text, engagement counts (likes, reposts, replies) and mentions, which give us the data needed for all four analyses: text analysis, hypothesis testing, clustering and network analysis.

## How the data was accessed and obtained

Data was collected from the Bluesky search API using the `atrrr` R package. We logged in with a Bluesky app password, which was typed at the prompt and never stored in the code or the GitHub repository.

The Bluesky code was run once on **1 Oct 2026**. It is commented out (`#`) below because running it again would return different posts and change our results. All outputs in this report come from the saved data files.

```{r}
# ---- Bluesky login and search function (run once on 1 Oct 2026, commented out) ----

# library(atrrr)
#
# # Log in to Bluesky (app password typed at the prompt, never saved in code)
# auth(user = "aakash18.bsky.social")
#
# # Search function that retries up to 3 times if Bluesky gives a server error (e.g. 502)
# get_posts <- function(t, limit = 500, tries = 3) {
#   for (i in 1:tries) {
#     p <- tryCatch(search_post(t, limit = limit), error = function(e) NULL)
#     if (!is.null(p)) return(p)
#     message("Retry ", i, " for: ", t)
#     Sys.sleep(5)
#   }
#   NULL
# }
```

The search function retries automatically because the Bluesky API occasionally returned "502 Bad Gateway" server errors during collection.

## R packages

- `atrrr`: connecting to Bluesky and collecting posts
- `dplyr`: working with tables (filtering, grouping, counting)
- `stringr`: working with text (detecting phrases, counting words)
- `knitr`: displaying tables in the report

```{r}
# Load the packages used in this section
library(dplyr)
library(stringr)
library(knitr)
```

## Choosing the topic (pilot)

Before committing to a topic, we collected about 500 posts for each of six candidate topics. We compared them on what our analyses need: enough posts, many different authors (for the network), engagement (for the hypothesis test) and mentions and replies (for the network).

```{r}
# ---- Pilot code (run once on 1 Oct 2026, commented out) ----

# topics <- c("nursing shortage", "AI healthcare", "housing crisis australia",
#             "climate change", "climate crisis", "net zero")
#
# # For each topic: collect posts and summarise volume, authors, engagement and mentions
# pilot <- lapply(topics, function(t) {
#   p <- get_posts(t)
#   Sys.sleep(2)
#   if (is.null(p)) return(tibble(topic = t, posts = NA))
#   tibble(
#     topic          = t,
#     posts          = nrow(p),
#     unique_authors = n_distinct(p$author_handle),
#     pct_liked      = round(mean(p$like_count > 0) * 100, 1),
#     median_likes   = median(p$like_count),
#     pct_mentions   = round(mean(str_detect(p$text, "@\\w")) * 100, 1),
#     pct_replied    = round(mean(p$reply_count > 0) * 100, 1)
#   )
# }) |> bind_rows()
#
# saveRDS(pilot, "pilot.rds")
```

```{r}
# Load and display the saved pilot results
pilot <- readRDS("pilot.rds")
kable(pilot, caption = "Pilot comparison of six candidate topics")
```

"Climate crisis" and "climate change" had the most unique authors (493 and 477), the highest share of liked posts (59.2% and 50.2%) and the highest mention rates (6.2% and 6.0%), so we chose climate. Comparing the two phrases also gave us a clear research angle: does the framing change the conversation? "AI healthcare" and "net zero" had a median of 0 likes, which would make engagement comparisons weak. "Nursing shortage" could not be collected because the API kept returning server errors.

## Full data collection

A check of 100 sample "climate crisis" posts showed some non-English posts and some posts that matched the search without using the phrase in their text. We therefore searched for the **exact phrases** (using quotes) with an **English-only filter** (`lang:en`), collecting up to 2,000 posts for each phrase.

```{r}
# ---- Main collection code (run once on 1 Oct 2026, 7:49pm AEST, commented out) ----

# # Quotes make Bluesky search for the exact phrase; lang:en keeps English posts only
# q_crisis <- get_posts('"climate crisis" lang:en', limit = 2000)
# Sys.sleep(5)
# q_change <- get_posts('"climate change" lang:en', limit = 2000)
#
# # Combine both searches and label where each post came from
# posts_raw <- bind_rows(
#   q_crisis |> mutate(search_term = "climate crisis"),
#   q_change |> mutate(search_term = "climate change")
# )
#
# saveRDS(posts_raw, "posts_raw.rds")
```

## Variables collected and data volume

```{r}
# Load the saved raw data
posts_raw <- readRDS("posts_raw.rds")

# Data volume: number of posts (rows) and variables (columns)
dim(posts_raw)
```

The raw data contains **4,148 posts and 22 variables**: 21 returned by the API, plus the `search_term` label we added.

```{r}
# Number of posts collected for each phrase
count(posts_raw, search_term)
```

The two searches returned a similar number of posts: 2,085 for "climate crisis" and 2,063 for "climate change".

```{r}
# Preview of the collected data (first 6 posts, main columns)
head(posts_raw[, c("author_handle", "search_term", "like_count", "reply_count")])
```

The preview shows the structure of the data: each row is one post, with its author, the phrase it was found with, and its engagement counts. Engagement is low for most posts (0 to 8 likes in this sample), which we explore further in the hypothesis test (2.2).

```{r}
# Example posts that matched the search but don't use the phrase in their text
# (these are removed during cleaning)
posts_raw |>
  filter(!str_detect(str_to_lower(text), "climate crisis|climate change")) |>
  select(search_term, text) |>
  head(3)
```

These posts were returned by the "climate crisis" search even though their text doesn't contain the phrase. They were likely matched through a linked article or quoted post. This shows why a cleaning step was needed to make sure each post actually uses the framing we are studying.

```{r}
# The main variables used in our analysis
variables <- data.frame(
  Variable = c("uri", "author_handle", "text", "like_count",
               "repost_count", "reply_count", "indexed_at", "search_term"),
  Description = c("Unique ID of the post",
                  "Bluesky username of the author",
                  "Text of the post",
                  "Number of likes",
                  "Number of reposts",
                  "Number of replies",
                  "Date and time the post was indexed",
                  "Which phrase the post was found with (added by us)"),
  Used_in = c("Cleaning (duplicates)", "Network", "Text analysis, clustering",
              "Hypothesis test", "Engagement summary", "Network",
              "Data description", "Cleaning")
)
kable(variables, caption = "Main variables used in the analysis")
```

The other columns returned by the API contain nested information about the author, the post (including reply links) and embedded content such as links and images. These are kept in `posts_raw.rds` for the network analysis.

## Cleaning the records

We cleaned the records in four steps:

1. Removed empty posts and posts that appeared in both searches.
2. Kept only posts whose text uses **exactly one** of the two phrases. Posts using neither phrase (matched only through a link or quoted post) or both phrases cannot be placed in one framing group.
3. Removed very short posts (fewer than 5 words), as they don't contain enough text to analyse.
4. Removed posts with identical text (bots and copy-paste spam).

```{r}
# Start from the saved raw data so the result is the same every time
n0 <- nrow(posts_raw)

# Step 1: remove empty posts and duplicates, then label each post's framing
posts_clean <- posts_raw |>
  filter(!is.na(text), text != "") |>
  distinct(uri, .keep_all = TRUE) |>
  mutate(
    text_lower = str_to_lower(text),
    has_crisis = str_detect(text_lower, "climate crisis"),
    has_change = str_detect(text_lower, "climate change"),
    framing = case_when(
      has_crisis & has_change ~ "both",
      has_crisis ~ "crisis",
      has_change ~ "change",
      TRUE ~ NA_character_
    ),
    n_words = str_count(text, "\\S+")
  )
n1 <- nrow(posts_clean)

# Step 2: keep posts that use only ONE of the two phrases
posts_clean <- posts_clean |> filter(!is.na(framing), framing != "both")
n2 <- nrow(posts_clean)

# Step 3: remove very short posts (fewer than 5 words)
posts_clean <- posts_clean |> filter(n_words >= 5)
n3 <- nrow(posts_clean)

# Step 4: remove posts with identical text (bots, spam)
posts_clean <- posts_clean |> distinct(text_lower, .keep_all = TRUE)
n4 <- nrow(posts_clean)
```

```{r}
# Number of posts left after each cleaning step
cleaning_steps <- data.frame(
  Step  = c("Raw posts", "Empty/duplicate posts removed",
            "Phrase in text (one framing only)", "5+ words", "Duplicate text removed"),
  Posts = c(n0, n1, n2, n3, n4)
)
kable(cleaning_steps, caption = "Posts remaining after each cleaning step")
```

Most steps removed only a small number of posts: 44 empty or duplicate posts, 13 very short posts and 94 posts with identical text. The biggest drop (4,104 to 2,507) came from removing posts that did not actually use either phrase in their text, as shown in the examples above.

```{r}
# Final number of clean posts in each framing
kable(count(posts_clean, framing), caption = "Clean posts by framing")
```

After cleaning, **2,400 posts** remain: 1,513 "climate change" and 887 "climate crisis". The "crisis" group lost more posts (2,085 collected, 887 kept) than the "change" group (2,063 collected, 1,513 kept), which suggests the "climate crisis" search matched more posts through links and quoted content. The clean data is saved as `posts_clean.rds` and used in 2.2, 2.3 and 2.4.

## Limitations

- **Snapshot in time:** the search returns recent posts, so our data only reflects the conversation around 1 Oct 2026. Results could be different at another time, for example during a major climate event.
- **API limits:** we collected up to 2,000 posts per phrase, not every post on the topic, and the API sometimes returned server errors (which is why "nursing shortage" failed in the pilot).
- **Search matching:** about 1,600 posts matched the search without using the phrase in their text, so they were removed. The English filter depends on the language tag set by the user's app, which may not always be correct.
- **Unequal groups:** there are more "change" posts (1,513) than "crisis" posts (887), so we compare proportions instead of raw counts where this matters.
- **Representativeness:** Bluesky users are not representative of all social media users or the general public, so our findings describe the Bluesky conversation only.

