---
title: "2.1 Data Collection"
author: "Aakash Jadhav"
date: "`r Sys.Date()`"
output: html_document
---

# 2.1 Data Collection

## How the data was accessed and obtained

Data was collected from the Bluesky search API using the `atrrr` R package. We logged in with a Bluesky app password, which was typed at the prompt and never stored in the code or the GitHub repository. The Bluesky code was run once on **1 Oct 2026**. It is commented out (`#`) below because running it again would return different posts and change our results. All outputs in this report come from the saved data files.

```{r}
# Bluesky login and search function (run once on 1 Oct 2026, commented out)
# library(atrrr) # connecting to Bluesky and collecting posts
# # Log in to Bluesky (app password typed at the prompt, never saved in code)
# auth(user = "aakash18.bsky.social")
# # Search function that retries up to 3 times if Bluesky gives a server error (e.g. 502)
# get_posts <- function(t, limit = 500, tries = 3) {
#   for (i in 1:tries) {
#     p <- tryCatch(search_post(t, limit = limit), error = function(e) NULL)
#     if (!is.null(p)) return(p)
#     message("Retry ", i, " for: ", t)
#     Sys.sleep(5)
#   }
#   NULL
# }
```

## R Packages used for analysis and visualisation

- `atrrr` was used only for data collection (the commented-out code above).

```{r}
library(dplyr)     # working with tables (filtering, grouping, counting)
library(stringr)   # working with text (detecting phrases, counting words)
library(knitr)     # displaying tables in the report
library(ggplot2)   # plots
library(tm)        # text mining
library(SnowballC) # stemming
library(wordcloud)    # word clouds  
library(RColorBrewer) # colour palettes 
library(igraph)     # network analysis
library(kableExtra) # fitting wide tables to the page width after converting it in to PDF format
```

## Choosing the topic (pilot)

Before committing to a topic, we collected a sample of about 550–600 posts for each of six candidate topics. We compared them on what our analyses need: enough posts, many different authors (for the network), engagement (for the hypothesis test) and mentions and replies (for the network).

```{r}
# Pilot code (run once on 1 Oct 2026, commented out) 
# topics <- c("nursing shortage", "AI healthcare", "housing crisis australia",
#             "climate change", "climate crisis", "net zero")
# # For each topic: collect posts and summarise volume, authors, engagement and mentions
# pilot <- lapply(topics, function(t) {
#   p <- get_posts(t)
#   Sys.sleep(2)
#   if (is.null(p)) return(tibble(topic = t, posts = NA))
#   tibble(
#     topic          = t,
#     posts          = nrow(p),
#     unique_authors = n_distinct(p$author_handle),
#     pct_liked      = round(mean(p$like_count > 0) * 100, 1),
#     median_likes   = median(p$like_count),
#     pct_mentions   = round(mean(str_detect(p$text, "@\\w")) * 100, 1),
#     pct_replied    = round(mean(p$reply_count > 0) * 100, 1)
#   )
# }) |> bind_rows()
#
# saveRDS(pilot, "pilot.rds")
```

```{r}
pilot <- readRDS("pilot.rds")    # Load and display the saved pilot results
kable(pilot, caption = "Pilot comparison of six candidate topics")
```

"Climate crisis" and "climate change" had the most unique authors (493 and 477) and the highest mention rates (6.2% and 6.0%), which matter most for the network analysis. "Climate crisis" also had the highest share of liked posts (59.2%). "Housing crisis australia" had good engagement (56.9% liked) but the fewest authors (359), which would give a smaller network. "AI healthcare" and "net zero" had a median of 0 likes, which would make engagement comparisons weak. "Net zero" had the highest reply rate (34.0%), but its lower engagement and fewer authors made climate the stronger overall choice. "Nursing shortage" could not be collected because the API kept returning server errors. We therefore chose climate, and comparing the two phrases gave us a clear research angle: whether the framing changes the conversation.

## Full data collection

A check of 100 sample "climate crisis" posts showed some non-English posts and some posts that matched the search without using the phrase in their text. We therefore searched for the **exact phrases** (using quotes) with an **English-only filter** (`lang:en`), requesting 2,000 posts for each phrase (the API returned slightly more: 2,085 and 2,063).

```{r}
# Main collection code (run once on 1 Oct 2026, 7:49pm AEST, commented out)
# # Quotes make Bluesky search for the exact phrase; lang:en keeps English posts only
# q_crisis <- get_posts('"climate crisis" lang:en', limit = 2000)
# Sys.sleep(5)
# q_change <- get_posts('"climate change" lang:en', limit = 2000)
# # Combine both searches and label where each post came from
# posts_raw <- bind_rows(
#   q_crisis |> mutate(search_term = "climate crisis"),
#   q_change |> mutate(search_term = "climate change")
# )
# saveRDS(posts_raw, "posts_raw.rds")
```

## Variables collected and data volume

```{r}
posts_raw <- readRDS("posts_raw.rds") # Load the saved raw data
dim(posts_raw)  # Data volume: number of posts (rows) and variables (columns)
```

The raw data contains **4,148 posts and 22 variables**: 21 returned by the API, plus the `search_term` label we added.

```{r}
# Number of posts collected for each phrase
count(posts_raw, search_term)
```

The two searches returned a similar number of posts: 2,085 for "climate crisis" and 2,063 for "climate change".

```{r}
# Preview of the collected data (first 6 posts, main columns)
head(posts_raw[, c("author_handle", "search_term", "like_count", "reply_count")])
```

The preview shows the structure of the data: each row is one post, with its author, the phrase it was found with, and its engagement counts. Engagement is low for most posts (0 to 8 likes in this sample), which we explore further in the hypothesis test (2.2).

```{r}
# Example posts that matched the search but don't use the phrase in their text
# (these are removed during cleaning)
posts_raw |>
  filter(!str_detect(str_to_lower(text), "climate crisis|climate change")) |>
  select(search_term, text) |>
  head(3)
```

These posts were returned by the "climate crisis" search even though their text doesn't contain the phrase. They were likely matched through a linked article or quoted post. This shows why a cleaning step was needed to make sure each post actually uses the framing we are studying.

```{r}
# Main variables collected and how they are used
variables <- data.frame(
  Variable = c("uri", "author_handle", "text", "like_count",
               "repost_count", "reply_count", "indexed_at", "search_term"),
  Description = c("Unique ID of the post",
                  "Bluesky username of the author",
                  "Text of the post",
                  "Number of likes",
                  "Number of reposts",
                  "Number of replies",
                  "Date and time the post was indexed",
                  "Which phrase the post was found with (added by us)"),
    Used_in = c("Cleaning (duplicates)", "Network labels (2.5)", "Text analysis, clustering (2.3, 2.4)",
              "Hypothesis test (2.2)", "Not used in analysis", "Pilot (topic choice)",
              "Not used in analysis", "Data volume (posts per phrase)")
)
kable(variables, caption = "Main variables collected and how they are used")
```

The other columns returned by the API include nested author, post and embed details (author_data, post_data, embed_data) and ready-made fields for replies (in_reply_to, in_reply_root), mentions, quotes, hashtags (tags), links and language (langs). These are kept in posts_raw.rds; the reply field (in_reply_to) provides the edges for the network analysis (2.5).

## Cleaning the records

```{r}
# Start from the saved raw data so the result is the same every time
n0 <- nrow(posts_raw)

# 4 steps used to clean the records
posts_clean <- posts_raw |>   # Step 1: remove empty posts and duplicates, then label each post's framing
  filter(!is.na(text), text != "") |>
  distinct(uri, .keep_all = TRUE) |>
  mutate(
    text_lower = str_to_lower(text),
    has_crisis = str_detect(text_lower, "climate crisis"),
    has_change = str_detect(text_lower, "climate change"),
    framing = case_when(
      has_crisis & has_change ~ "both",
      has_crisis ~ "crisis",
      has_change ~ "change",
      TRUE ~ NA_character_
    ),
    n_words = str_count(text, "\\S+")
  )
n1 <- nrow(posts_clean)
# Step 2: keep posts that use only ONE of the two phrases
posts_clean <- posts_clean |> filter(!is.na(framing), framing != "both")
n2 <- nrow(posts_clean)

posts_clean <- posts_clean |> filter(n_words >= 5) # Step 3: remove very short posts (fewer than 5 words)
n3 <- nrow(posts_clean)

posts_clean <- posts_clean |> distinct(text_lower, .keep_all = TRUE) 
n4 <- nrow(posts_clean)
# saveRDS(posts_clean, "posts_clean.rds")   # saved for 2.2, 2.3 and 2.4
```

```{r}
# Number of posts left after each cleaning step
cleaning_steps <- data.frame(
  Step  = c("Raw posts", "Empty/duplicate posts removed",
            "Phrase in text (one framing only)", "5+ words", "Duplicate text removed"),
  Posts = c(n0, n1, n2, n3, n4)
)
kable(cleaning_steps, caption = "Posts remaining after each cleaning step")
```

Most steps removed only a small number of posts: 44 empty or duplicate posts, 13 very short posts and 94 posts with identical text. The biggest drop (4,104 to 2,507) came from removing posts that did not actually use either phrase in their text, as shown in the examples above.

```{r}
# Final number of clean posts in each framing
kable(count(posts_clean, framing), caption = "Clean posts by framing")
```

After cleaning, **2,400 posts** remain: 1,513 "climate change" and 887 "climate crisis". Only 887 crisis-framed posts remain from the 2,085 results of the "climate crisis" search, compared with 1,513 change-framed posts from the 2,063 results of the "climate change" search. This suggests the "climate crisis" search matched more posts through links and quoted content. The clean data is saved as `posts_clean.rds` and used in 2.2, 2.3 and 2.4.

*Limitations*
The main limitations of the data collection (a single snapshot in time, API restrictions, search matching and representativeness) are discussed in Section 3. A further point is that the two framing groups are unequal in size, with 1,513 "change" posts and 887 "crisis" posts, so proportions rather than raw counts are compared wherever this affects the results.
