---
title: "2.3 Text/Content Analysis and Visualisation"
author: "Pushti Patel & Aakash Jadhav"
date: "2026-10-03"
output: html_document
---

```{r, message=FALSE, warning=FALSE}
# Packages
library(dplyr); library(stringr); library(ggplot2); library(knitr)
library(tm); library(SnowballC); library(wordcloud); library(RColorBrewer)
```

# 2.3 Text/Content Analysis and Visualisation

**Research question:** Do "climate crisis" and "climate change" posts differ in the words and topics they use?

**What?**
We explore the language used in the 2,400 clean posts: which words appear most often overall, and which words are most common in "climate crisis" posts compared with "climate change" posts.

**Why?**
The hypothesis test (2.2) found no evidence that engagement depends on framing. Text analysis lets us check whether the two framings differ in **what** people talk about, even if they get similar engagement.

**Missing and unsuitable records:**
Empty posts, duplicates, very short posts and posts without either phrase were already removed in 2.1, leaving 2,400 posts. This section prepares the **text** inside those posts for word analysis; the original posts are not changed.

```{r}
# Load the cleaned posts from 2.1
posts <- readRDS("posts_clean.rds")

# Check for missing text
sum(is.na(posts$text))
```

There is no missing text (0 posts), so all 2,400 posts can be analysed.

## Text summaries

```{r}
posts |>    # Post length (number of words) by framing
  group_by(framing) |>
  summarise(posts = n(),
            mean_words = round(mean(n_words), 1),
            median_words = median(n_words),
            max_words = max(n_words))
```

Posts in both framings are a similar length. "Crisis" posts average 34.1 words (median 36) and "change" posts 33.1 words (median 35). The maximum of 61–64 words reflects Bluesky's short post format. Any difference between the framings is therefore about *what* is said, not how much.

## Cleaning the text

Before counting words, we remove things that would show up as meaningless "words":

- **Links and @usernames**, otherwise punctuation removal turns them into junk like "httpswwwtheguardiancom"
- **Accented letters:** "ñ" is converted to "n" first, so "El Niño" (a climate pattern that appears often) is kept as "nino" instead of being cut to "nio"
- **Emojis and special characters**
- **Numbers and punctuation**
- **Stopwords** (common words such as "the", "and", "is")
- **The search words** "climate", "crisis", "change" and their hashtag versions (#ClimateChange, #ClimateCrisis): every post contains them by design, so they would dominate without telling us anything
- **Generic filler words** such as "will", "like", "just", "can" and "get", which our first run showed dominated the top words without carrying meaning
- **"amp"** (left over from the "&" symbol) and **"bigearthdataai"** (an account that appears to post repeated automated content)

We then **stem** the words (e.g. "warming" and "warmed" become "warm") so different forms of the same word are counted together. Stemmed words therefore appear in shortened form in the outputs, e.g. "peopl" (people), "energi" (energy), "secur" (security) and "futur" (future).

```{r}
# Remove links and @usernames, convert ñ to n (so "El Niño" is kept as "nino")& remove emojis/special characters
clean_text <- posts$text |>
  str_remove_all("http\\S+|www\\.\\S+") |>
  str_remove_all("@\\S+") |>
  str_replace_all("[ñÑ]", "n") |>
  iconv(to = "ASCII", sub = "")
```

```{r}
# Words to remove: stopwords, the search words (and hashtag versions),
# generic filler words that carry no meaning, and an automated account name
custom_stop <- c(stopwords("english"),
                 "climate", "crisis", "change", "climatechange", "climatecrisis", "amp",
                 "will", "like", "just", "can", "get", "one", "dont", "now", "also",
                 "even", "much", "way", "see", "know", "thing", "things", "want",
                 "let", "use", "still", "really", "that", "isnt", "doesnt", "look",
                 "come", "many", "lot", "well", "got", "going", "new", "make",
                 "say", "says", "said", "bigearthdataai")

# Build the corpus (one document per post) and clean it
corpus <- VCorpus(VectorSource(clean_text))
corpus <- tm_map(corpus, content_transformer(tolower))   # lowercase
corpus <- tm_map(corpus, removeNumbers)                  # remove numbers
corpus <- tm_map(corpus, removePunctuation)              # remove punctuation
corpus <- tm_map(corpus, removeWords, custom_stop)       # remove stopwords and custom words
corpus <- tm_map(corpus, stemDocument)                   # stem words
corpus <- tm_map(corpus, stripWhitespace)                # remove extra spaces
```

**Example: one post before cleaning**

```{r}
posts$text[1] # Original text of the first post
```

**The same post after cleaning**

```{r}
as.character(corpus[[1]]) # Cleaned text of the first post
```

The example shows the effect of cleaning: punctuation, stopwords ("are", "not", "a"), filler words ("just", "get") and the phrase "climate crisis" are removed, and words are stemmed (e.g. "party" → "parti", "permanently" → "perman", "promising" → "promis"). One limitation is visible: filler words are removed before stemming, so inflected forms such as "using" still appear after stemming as "use".

## Representing the text

We represent the text as a **term-document matrix**: each row is a word, each column is a post, and each cell counts how many times that word appears in that post (vector space model). In 2.4 this matrix is weighted with TF-IDF for clustering.

```{r}
tdm <- TermDocumentMatrix(corpus) # Term-document matrix
tdm
```

The term-document matrix contains 6,878 unique terms across 2,400 posts. Only 35,974 of its 16,507,200 cells are non-zero, so the matrix is over 99.7% sparse, and each post contains about 15 distinct terms on average. This is typical of short social media posts: most words appear in only a few posts, which is why TF-IDF weighting is used for clustering in 2.4.

```{r}
# Convert to a matrix and count how often each word appears overall
m <- as.matrix(tdm)
word_freq <- sort(rowSums(m), decreasing = TRUE)
freq_df <- data.frame(word = names(word_freq), freq = word_freq)

sum(colSums(m) == 0) # Check for posts with no words left after cleaning
```

Two posts contained no words after cleaning (likely because their text consisted only of stopwords, links or mentions). They are kept in the data but have no effect on word frequencies.

## Visualisation 1: Most frequent words

```{r}
ggplot(head(freq_df, 20), aes(x = reorder(word, freq), y = freq)) +    # Bar chart of the 20 most frequent words
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 20 Most Frequent Words (stemmed)",
       x = "Word", y = "Frequency") +
  theme_minimal()
```

After removing filler words, the most frequent terms point to several themes: **human impact** ("peopl", "human", "impact", "caus"), **energy** ("fuel", "fossil", "energi"), **politics and security** ("trump", "nation", "secur", "report"), **urgency** ("need", "time", "action") and **extreme weather** ("nino", for El Niño, which appears 146 times). The conversation is therefore not only about climate science, but also about energy, politics and people's lives.

## Visualisation 2: Word cloud of all posts

```{r, fig.width = 6.5, fig.height = 4.5}
set.seed(3020)                  # Word cloud of the 150 most frequent words (bigger = more frequent)
wordcloud(words = freq_df$word, freq = freq_df$freq,
          max.words = 150, random.order = FALSE,
          colors = brewer.pal(8, "Dark2"))
```

The word cloud shows the same dominant themes. Beyond the top 20, smaller words show the wider range of the discussion: extreme weather ("flood", "heat", "water"), energy sources ("oil", "gas"), effects on society ("health", "food", "threat", "war") and government action ("govern", "address", "stop"). The word "hoax" also appears, suggesting some posts question whether climate change is real.

## Comparing "crisis" and "change" posts

To compare the framings, we count word frequencies separately for "crisis" posts and "change" posts.

```{r}
# Word frequencies within each framing
# (TRUE = all rows/words; the second part picks the posts in each framing)
crisis_freq <- sort(rowSums(m[TRUE, posts$framing == "crisis"]), decreasing = TRUE)
change_freq <- sort(rowSums(m[TRUE, posts$framing == "change"]), decreasing = TRUE)

# Top 10 words in each framing, as a % of all words in that framing
# (percentages make the groups comparable, since there are more "change" posts)
top_crisis <- data.frame(crisis_word = names(crisis_freq)[1:10],
                         crisis_pct = round(crisis_freq[1:10] / sum(crisis_freq) * 100, 2))
top_change <- data.frame(change_word = names(change_freq)[1:10],
                         change_pct = round(change_freq[1:10] / sum(change_freq) * 100, 2))

kable(cbind(top_crisis, top_change), row.names = FALSE,
      col.names = c("Crisis word", "% of crisis words", "Change word", "% of change words"),
      caption = "Top 10 words in each framing")
```

Five words appear in both top-10 lists ("world", "peopl", "fuel", "global", "year"), so the two framings share a common core. The distinctive words differ, though. **"Crisis" posts** feature "nation", "secur", "report", "nino" and "time", which suggests news reports, national security and extreme weather. **"Change" posts** feature "think", "need", "human", "impact" and "trump", which suggests opinions, political debate and human consequences.

```{r}
# Posts mentioning El Niño in each framing
# (matches "niño" or "nino" as a whole word, so names like "Antonino" are not counted)
posts |>
  mutate(mentions_nino = str_detect(str_to_lower(text), "\\bni[ñn]o\\b")) |>
  group_by(framing) |>
  summarise(posts = n(),
            nino_posts = sum(mentions_nino),
            pct = round(mean(mentions_nino) * 100, 1))
```

Posts using "Niño" or "Nino" as a separate word appear in 68 "crisis" posts (7.7% of 887) and 58 "change" posts (3.8% of 1,513). "Crisis" posts were therefore about twice as likely to mention El Niño, which suggests the "crisis" framing is more often used when discussing extreme weather events. (Hashtags such as #ElNiño, where the words are joined, are not included in this count.)

## Visualisation 3 and 4: Word clouds by framing

### Climate crisis posts

```{r, fig.width = 6.5, fig.height = 4.5}
set.seed(3020)       # Word cloud of "crisis" posts
wordcloud(words = names(crisis_freq), freq = crisis_freq,
          max.words = 100, random.order = FALSE,
          colors = brewer.pal(8, "Dark2"))
```

### Climate change posts

```{r, fig.width = 6.5, fig.height = 4.5}
set.seed(3020)      # Word cloud of "change" posts
wordcloud(words = names(change_freq), freq = change_freq,
          max.words = 100, random.order = FALSE,
          colors = brewer.pal(8, "Dark2"))
```

The clouds confirm the table. "peopl", "world", "year" and "fuel" are prominent in both. The "crisis" cloud gives much more space to "nation", "report", "secur" and "nino", along with urgent words such as "death", "war" and "warn", while "trump" does not appear in it. In the "change" cloud, after "peopl" and "year", the next largest words include "think", "need", "human" and "trump", and "hoax" and "real" also stand out; "hoax" does not appear in the "crisis" cloud. Political words such as "vote" and "polici" also appear only in the "change" cloud. This suggests the "change" framing is more often used in debates about politics and whether climate change is real, while "crisis" is more often used alongside news reports, national security and extreme weather.
