---
title: "2.4 Clustering"
author: "Aakash Jadhav"
date: "`r Sys.Date()`"
output: html_document
---
```{r, message=FALSE}
library(dplyr); library(stringr); library(knitr); library(ggplot2); library(tm); library(SnowballC)

# Rebuild the cleaned posts and text corpus exactly as in 2.3
posts <- readRDS("posts_clean.rds")
clean_text <- posts$text |>
  str_remove_all("http\\S+|www\\.\\S+") |>
  str_remove_all("@\\S+") |>
  str_replace_all("[ñÑ]", "n") |>
  iconv(to = "ASCII", sub = "")
custom_stop <- c(stopwords("english"),
                 "climate", "crisis", "change", "climatechange", "climatecrisis", "amp",
                 "will", "like", "just", "can", "get", "one", "dont", "now", "also",
                 "even", "much", "way", "see", "know", "thing", "things", "want",
                 "let", "use", "still", "really", "that", "isnt", "doesnt", "look",
                 "come", "many", "lot", "well", "got", "going", "new", "make",
                 "say", "says", "said", "bigearthdataai")
corpus <- VCorpus(VectorSource(clean_text))
corpus <- tm_map(corpus, content_transformer(tolower))
corpus <- tm_map(corpus, removeNumbers)
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeWords, custom_stop)
corpus <- tm_map(corpus, stemDocument)
corpus <- tm_map(corpus, stripWhitespace)
```

# 2.4 Clustering

**Research question:** Can posts be grouped into distinct themes based on their words, and are the two framings distributed differently across these themes?

**What?**
We investigate whether the 2,400 posts form meaningful groups (clusters) based on the words they use, and what theme each cluster represents.

**Why?**
The text analysis (2.3) suggested several overlapping themes (energy, politics and security, human impact, extreme weather, debate over whether climate change is real). Clustering identifies these themes directly from the posts, without using the framing labels. We can then check whether "crisis" and "change" posts fall into different themes.

## Representation: document-term matrix and TF-IDF

We reuse the cleaned posts and text (`corpus`) from 2.3, so both sections analyse exactly the same words.

```{r}
dtm <- DocumentTermMatrix(corpus) # Document-term matrix: one row per post, one column per word
dtm
```

The document-term matrix has 2,400 posts and 6,878 terms, the same vocabulary as in 2.3. Most of these terms appear in only one or two posts. Rare words cannot help group posts together, and they add noise to the distances, so we keep only terms that appear in at least 0.5% of posts (about 12 or more posts).

```{r}
dtm_small <- removeSparseTerms(dtm, 0.995)   # Keep only terms that appear in at least 0.5% of posts
dtm_small
```

This reduces the vocabulary from 6,878 to 667 terms. The matrix is still very sparse: only 21,331 of its 1,600,800 cells (1.3%) are non-zero.

```{r}
# Posts with none of the remaining terms cannot be compared, so they are removed
tf <- as.matrix(dtm_small)
keep <- rowSums(tf) > 0
sum(!keep)
```

15 posts contained none of the 667 remaining terms, so they are removed, leaving 2,385 posts for clustering.

We weight the matrix with **TF-IDF**. It gives more weight to words that are frequent in a post but rare across all posts, so distinctive words count more than words that appear everywhere. tm's TF-IDF also adjusts for post length, so longer posts don't dominate.

```{r}
# TF-IDF weighting, then keep only the posts with at least one remaining term
X <- as.matrix(weightTfIdf(dtm_small))[keep, TRUE]
posts_k <- posts[keep, ]
dim(X)
```

## Distance measure: cosine distance

We use **cosine distance**. It compares posts by the *direction* of their word vectors, i.e. which words they use, regardless of length. Euclidean distance would treat two posts as different simply because one is longer. Cosine distance is 0 for posts using the same words in the same proportions, and 1 for posts with no words in common.

```{r}
# Cosine distance = 1 - cosine similarity
X_norm <- X / sqrt(rowSums(X^2))            # scale each post's vector to length 1
cos_dist <- as.dist(1 - X_norm %*% t(X_norm))
```

## First attempt: hierarchical clustering

We first tried **hierarchical clustering**, because it works directly on the cosine distance matrix. We compared two linkage methods:

- **Complete linkage:** the distance between clusters is the *largest* distance between their posts
- **Group average linkage:** the distance between clusters is the *average* distance between their posts

```{r}
# Hierarchical clustering with two linkage methods
hc_complete <- hclust(cos_dist, method = "complete")
hc_average  <- hclust(cos_dist, method = "average")
```

For each linkage, we cut the tree into 2 to 8 clusters and compared the cluster sizes. Useful clusters should be reasonably balanced: a solution with one huge cluster and several tiny ones does not summarise the data well.

```{r}
# Cluster sizes (largest to smallest) for k = 2 to 8, for each linkage method
sizes <- function(hc, k) paste(sort(table(cutree(hc, k)), decreasing = TRUE), collapse = " / ")

size_table <- data.frame(
  k = 2:8,
  complete = sapply(2:8, function(k) sizes(hc_complete, k)),
  average  = sapply(2:8, function(k) sizes(hc_average, k))
)
kable(size_table, caption = "Hierarchical clustering: cluster sizes for each k and linkage method")
```

Neither linkage produces useful clusters. **Complete linkage** keeps at least 2,363 of the 2,385 posts in one cluster for every k, and only splits off tiny groups of 2–5 posts. **Average linkage** is better, but one cluster still holds at least 2,019 posts (85%) at every k.

```{r, fig.width = 6.5, fig.height = 4.5}
# Dendrogram of the average linkage tree
plot(hc_average, labels = FALSE, hang = -1,
     main = "Dendrogram (cosine distance, average linkage)",
     xlab = "Posts", sub = "")
```

The dendrogram confirms this. Almost all merges happen at heights between about 0.8 and 1.0, meaning most posts are joined to the tree only when clusters are already very dissimilar. Only a few narrow branches reach lower heights; these are small groups of near-identical posts, which form the tiny clusters in the table above.

The reason is the data itself. After removing rare terms, each post contains only about 9 distinct terms out of 667, so many pairs of posts share no words at all and have a cosine distance close to 1. With so many distances close to 1, hierarchical clustering has little structure to work with: it merges a few near-identical posts early and leaves almost everything else in one large group.

## Final method: k-means on normalised TF-IDF vectors

We therefore use **k-means** on the TF-IDF vectors after scaling each post's vector to length 1 (`X_norm`). For vectors of length 1, the squared Euclidean distance used by k-means equals twice the cosine distance, so k-means on `X_norm` still groups posts by cosine similarity, the same distance measure as above. Unlike hierarchical clustering, k-means builds clusters around central points (centroids), which tends to give more balanced groups. We use 20 random starts (`nstart = 20`) to avoid a poor solution.

## Choosing the number of clusters

We run k-means for k = 2 to 10 and record the **total within-cluster sum of squares** (how spread out the posts are within their clusters). This always decreases as k increases. We look for an **elbow**, where adding more clusters stops giving a large improvement. We also check the cluster sizes for balance.

```{r}
# Run k-means for k = 2 to 10 (same random seed for each, so results are reproducible)
fits <- lapply(2:10, function(k) {
  set.seed(3020)
  kmeans(X_norm, centers = k, nstart = 20, iter.max = 50)
})
```

```{r}
# Elbow plot: total within-cluster sum of squares for each k
wss <- sapply(fits, function(f) f$tot.withinss)
plot(2:10, wss, type = "b", pch = 19,
     main = "Elbow plot (k-means on normalised TF-IDF)",
     xlab = "Number of clusters (k)", ylab = "Total within-cluster sum of squares")
```

The total within-cluster sum of squares decreases steadily and almost linearly from k = 2 to k = 10, with no clear elbow, and the overall reduction is small relative to the total. This shows that the posts do not divide into strongly separated groups, which is consistent with how sparse short posts are. Since the elbow method does not identify a single best k, we choose k using cluster balance and how interpretable the clusters are.

```{r}
# Cluster sizes (largest to smallest) for each k
kable(data.frame(k = 2:10,
                 sizes = sapply(fits, function(f) paste(sort(f$size, decreasing = TRUE), collapse = " / "))),
      caption = "k-means: cluster sizes for each k")
```

We chose k = 5. Unlike hierarchical clustering, k-means separates several medium-sized groups from the main body of posts: at k = 5 the clusters contain 1,974, 197, 110, 56 and 48 posts. Each of the four smaller clusters has a distinct and interpretable theme (see below). Larger values of k give more evenly sized clusters, but with no elbow there is no statistical reason to prefer them, and more clusters are harder to interpret and present. The choice of k is therefore a judgement, which we note as a limitation.

```{r}
# Chosen number of clusters (decided from the elbow plot and size table above)
k <- 5
clusters <- fits[[k - 1]]$cluster
table(clusters)
```

## Cluster themes

To identify each cluster's theme, we find the words with the highest average weight within the cluster: the words most characteristic of its posts.

```{r}
# Top 8 most characteristic words in each cluster (highest average normalised TF-IDF)
top_terms <- sapply(1:k, function(j) {
  avg <- colMeans(X_norm[clusters == j, TRUE, drop = FALSE])
  paste(names(sort(avg, decreasing = TRUE))[1:8], collapse = ", ")
})

kable(data.frame(cluster = 1:k,
                 posts = as.vector(table(clusters)),
                 top_words = top_terms),
      caption = "Size and most characteristic words of each cluster")
```

**Based on their most characteristic words, the five clusters can be described as**:

- Cluster 1 (197 posts), fossil fuels, industry and legal cases: "fuel", "fossil", "industri", "compani", together with "justic", "alito" and "case", which suggest discussion of a court case involving fossil fuel companies, and "hoax", which suggests scepticism.
- Cluster 2 (56 posts), national security report: "secur", "nation", "suppress", "starmer" and "report" suggest a UK news story about a suppressed national security report, involving Prime Minister Keir Starmer.
- Cluster 3 (110 posts), opinions and solutions: "think", "peopl", "solv", "work", "start" suggest personal views on what should be done.
- Cluster 4 (48 posts), extreme weather and deaths: "death", "nino", "toll", "project", "temperatur" suggest reports of projected death tolls linked to El Niño and rising temperatures.
- Cluster 5 (1,974 posts, 82.8%), general climate discussion: broad words such as "peopl", "world", "need", "real", "global" and "impact", the same top words as the whole dataset in 2.3.

## Visualising the clusters (MDS)

We visualise the clusters with **multidimensional scaling**. MDS places the posts in two dimensions so that the distances between points approximate their cosine distances. Unlike PCA, it works with any distance measure.

```{r}
# Classical MDS on the cosine distances, reduced to 2 dimensions
mds <- cmdscale(cos_dist, k = 2)
mds_df <- data.frame(dim1 = mds[, 1], dim2 = mds[, 2],
                     cluster = factor(clusters),
                     framing = posts_k$framing)

ggplot(mds_df, aes(x = dim1, y = dim2, colour = cluster)) +
  geom_point(alpha = 0.6, size = 1.2) +
  labs(title = "Posts in 2D (MDS on cosine distance), coloured by cluster",
       x = "MDS dimension 1", y = "MDS dimension 2", colour = "Cluster") +
  theme_minimal()
```

The MDS plot shows a dense core containing most posts, where clusters 5 (general discussion), 1 (fossil fuels and legal cases) and 3 (opinions and solutions) overlap. Two clusters are clearly separated. Cluster 2 (national security report) forms a distinct arm towards the upper right, and cluster 4 (extreme weather and deaths) forms a distinct arm towards the bottom. These two clusters use vocabulary that is very different from the rest of the posts. The overlap in the core is partly because MDS shows only two dimensions of the cosine distances, so clusters that overlap here may still be separated in the full space.


## Linking clusters to framing

```{r}
# Number of posts of each framing in each cluster
tab_cf <- table(cluster = clusters, framing = posts_k$framing)
tab_cf
```

```{r}
round(prop.table(tab_cf, 1), 3)   # Proportion of "change" and "crisis" posts within each cluster
```

```{r}
# Stacked bar chart: framing mix within each cluster
ggplot(mds_df, aes(x = cluster, fill = framing)) +
  geom_bar(position = "fill") +
  geom_hline(yintercept = mean(posts_k$framing == "crisis"), linetype = "dashed") +
  scale_fill_manual(values = c(change = "steelblue", crisis = "firebrick")) +
  labs(title = "Framing mix within each cluster",
       subtitle = "Dashed line = overall share of crisis posts",
       x = "Cluster", y = "Proportion of posts", fill = "Framing") +
  theme_minimal()
```

Overall, 37.0% of the clustered posts use "crisis" framing (dashed line). Clusters 2 and 4 are dominated by "crisis" posts: 78.6% of the national security cluster and 89.6% of the extreme weather cluster. In contrast, "crisis" is under-represented in clusters 1 (31.0%) and 3 (26.4%), the fossil fuel and opinion clusters. The general discussion cluster (35.7%) is close to the overall share.

To check whether the framing mix really differs between clusters, we use the chi-squared test of independence, as in 2.2.

```{r}
# Chi-squared test: are cluster and framing independent?
test_cf <- chisq.test(tab_cf)
round(test_cf$expected, 1)   # expected counts should be at least 5
```

```{r}
test_cf
```

H0: cluster and framing are independent. H1: cluster and framing are not independent. All expected counts are above 5 (the smallest is 17.8), so the test is valid. The chi-squared test gives X² = 108.29 (df = 4), p < 0.001, so we reject H0: the framing mix differs significantly between clusters. "Crisis" framing is concentrated in posts about specific news events (the security report and El Niño deaths), while "change" framing is more common in opinion and debate posts.

## Linking clusters to engagement

```{r}
# Engagement in each cluster
posts_k |>
  mutate(cluster = factor(clusters)) |>
  group_by(cluster) |>
  summarise(posts = n(),
            median_likes = median(like_count),
            pct_5plus_likes = round(mean(like_count >= 5) * 100, 1))
```

Every cluster has a median of 1 like. The share of posts with 5 or more likes ranges from 19.6% (cluster 2) to 27.3% (cluster 3), so engagement is broadly similar across themes. Clusters 3 and 4 are slightly higher, but they are small (110 and 48 posts), so these differences may not be meaningful. This is consistent with 2.2: neither framing nor theme is strongly linked to engagement.

*Limitations*: the clusters are only weakly separated (no clear elbow, one large general cluster), the choice of k = 5 is a judgement, and the two smallest clusters may each reflect a single news story rather than a lasting theme.

